### 2017-06-20 latex/paper.tex
first commit of latex

% @Author: YangZhou
% @Date:   2017-06-20 20:27:38
% @Last Modified by:   YangZhou
% @Last Modified time: 2017-06-20 22:10:07
\documentclass[%
 reprint,
 amsmath,amssymb,
 aps,
prb,
]{revtex4-1}
\preprint{APS/123-QED}
\begin{document}
\title{Ultra low thermal conductivity of quasi 1D bulk material β-Bi4I4}
\author{Ann Author}
 \altaffiliation[Also at ]{Physics Department, XYZ University.}%Lines break automatically or can be forced with \\
\author{Second Author}%
 \email{Second.Author@institution.edu}
\affiliation{%
 Authors' institution and/or address\\
 This line break forced with \textbackslash\textbackslash
}%
\begin{abstract}
An article usually includes an abstract, a concise summary of the work
covered at length in the main body of the article. 
\begin{description}
\item[Keywords]
Secondary publications and information retrieval purposes.
\end{description}
\end{abstract}

\maketitle

hello, world % This is comment
\end{document}